---
title: 'Assignment 4: Predicting Stock Returns'
subtitle: 'Exploratory Data Analysis (1/3)'
author: "Jesús Calderón"
date: "02/12/2021"
output: 
  html_document:
    toc: FALSE
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```




## Exploratory Data Analysis

+ Describe the data. You can find a description of the data in the Introduction (pages 2-3) of James et al, [Introduction to Statistical Learning](https://www.statlearning.com/). (As well, you can find some interesting discussion on modelling it in Chapter 4).
+ Display the data using pair plots or other forms of graphical device. 
+ Calculate and show descriptive statistics: include, at least, a measure of centrality (mean or median) and a measure of dispersion (standard deviation or IQR), as well as important quantiles and extremes.

> Guidance: minimum requirements are:
> 
> + Brief description of the data, including a description of the variables, number of observations, and date range.
> + A set of charts showing the potential relationship between predictors and target variable. Charts showing the relationship among predictors are a good addition, but the report should focus on predicting the target variables.
> + Summary statistics in general should be produced with `group_by()` and `summarise()` operations. Some submissions will use `summary()`, but this function does not show dispersion metrics.
>
> As a general observation: correlations are low across predictors and there does not appear to be a discernible relationship between the target variable and any of the predictors.

```{r}
library(tidyverse)
library(ISLR)
library(GGally)
library(knitr)
```

```{r}
data(Smarket)
market_dt <- Smarket %>%
  as_tibble() %>%
  select(-Today) %>%
  mutate(Direction = fct_relevel(Direction, 'Up'))
```

```{r}
market_dt %>%
  select(-Year) %>%
  ggpairs(aes(fill = Direction), progress = FALSE) +
  scale_fill_brewer(type = 'qual', palette = 2) +
  theme_minimal() +
  labs(title = 'Exploratory Data Analysis', 
       subtitle = 'Pair Plots and Correlations')
```


```{r}
market_dt %>%
  select(-Year) %>%
  pivot_longer(-Direction) %>%
  group_by(name) %>%
  summarise(min = min(value),
            q25 = quantile(value, 0.25),
            median = median(value),
            average = mean(value),
            q75 = quantile(value, 0.75),
            max = max(value),
            stdev = sd(value),
            iqr = IQR(value)) %>%
  kable(digits = 3, 
        caption = 'Summary statistics')
```


```{r}
market_dt %>%
  select(-Year) %>%
  pivot_longer(-Direction) %>%
  group_by(name, Direction) %>%
  summarise(min = min(value),
            q25 = quantile(value, 0.25),
            median = median(value),
            average = mean(value),
            q75 = quantile(value, 0.75),
            max = max(value),
            stdev = sd(value),
            iqr = IQR(value)) %>%
  kable(digits = 3, 
        caption = 'Summary statistics grouped by Direction')
```